Constraints on magic state protocols from the statistical mechanics of \\Wigner negativity

### 2021-06-28 magic_paper.tex
Potentially my final edits. Refs to be added.

\title{General constraints on magic state protocols from the statistical mechanics of \\Wigner negativity}

@@

To address this, concrete distillation protocols have been developed, such as the Bravyi-Haah qubit protocol that provides a quadratic reduction in noise per cycle~\cite{Bravyi2012}. Such distillation rates have been improved in more recent works~\cite{Jones_2013, haah2017magic, Hastings2018, Litinski_2019} and protocols for qudits of odd dimension have also been proposed~\cite{CampbellAnwar_2012, Anwar_2012, Dawkins_2015, Krishna2019, cit:prakash}. There is also analysis of magic protocols from the perspective of magic monotones, which provide upper bounds on distillation rates and find application in the analysis of gate synthesis~\cite{Campbell_2017, Howard_2017, Prakash_2018, Seddon_2021,leone2021renyi}. These frameworks for magic view magic states as resource states with respect to a natural class of quantum operations that are considered cheap, or free, such as stabilizer operations~\cite{Gour_2019, cit:ahmadi, cit:seddon, Wang_2019}. Beyond quantum computing, recent work has also looked at how magic can be used to analyse many-body physics~\cite{Sarkar_2020} and conformal field theories~\cite{White_2021} as well as how it is connected with contextuality~\cite{Vega_2017, cit:howard2, Zurel_2020, Okay_2021}.

Recent work towards fault-tolerance is increasingly bridging the gap between abstract theory and experiment. Extensive work has been done on error mitigation~\cite{jones_2012, Li_2017, Temme_2017, Endo_2018, McClean_2017}, and the incorporation of hardware physics into the theoretical modes~\cite{Kandala_2019, holmes_resource_2019, Colless_2018, song2018quantum, Bravyi_2021}. For example, the XZZX code~\cite{bonilla_ataides_xzzx_2021} is a variant of the surface code that incorporates noise bias explicitly and has been shown to attain the hashing bound of random codes~\cite{Bennett_1996}.

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated in the distillation bounds. This is achieved by considering how a given magic state protocol transforms a \emph{pair} of quantum states -- one being a noisy magic state, and the other a stabilizer state that is distinguished by the physics of the system (for example a state at a characteristic temperature, or the maximally mixed state) and acts as a reference state for the protocol. The magic distillation bounds can then be expressed in terms of the physics of the reference state (e.g.~free energy changes in the case of temperature). Such bounds are therefore of interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols~\cite{Tuckett_2019, Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in key ways. We obtain distillation upper bounds without the use of monotones, but instead use tools from statistical mechanics and recent work in single-shot resource theories~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. Our analysis relies on the discrete Wigner representation of quantum systems, in which all states and operations can be described on a discrete phase space~\cite{Ferrie_2008, Okay_2021}. Crucially, we focus on magic states with negativity in their Wigner representation, which is known to be a necessary condition for universality in the state-injection model~\cite{cit:veitch, cit:mari, cit:gottesman, cit:knill, Campbell_2011}. However, taking a statistical mechanical perspective in this context raises a problem: the standard Boltzmann entropy is not defined for quasi-distributions. We circumvent this obstacle by making use of a more fundamental tool -- majorization theory~\cite{cit:marshall, Veinott_1971, Ruch_1976}. To our knowledge majorization of quasi-distributions has not been considered in quantum physics before, and therefore our analysis is also of relevance to other studies of non-classicality in quantum systems~\cite{Fine_1982, Allahverdyan_2018, arvidsson_2020, halpern_2018, Lostaglio_2018, Levy_2020}.

@@

We find that the analysis in the presence of negativity is non-trivial, displays a range of features that do not appear in classical statistical mechanics, and leads to a picture of Wigner negativity in a quantum circuit being described as non-classical free energy that is processed under stochastic dynamics. This is demonstrated by Lorenz curves having a non-trivial maximum that is related to the mana of a state (Section~\ref{sec:lc}), and by showing that single-shot entropies for general Wigner quasi-distributions are well-defined, obey data-processing under stabilizer operations and can take on negative values (Section~\ref{section:entropies}).

@@

We show that the mana of a quantum state also arises as a residue in the $\alpha$--parameter for R\'{e}nyi entropies, and gives the rate of divergence to $-\infty$ as we approach the Shannon entropy of a Wigner distribution (Equation (\ref{eq:mana-divergence})).
Finally, in Section~\ref{sec:lower_bounds} we discuss how our approach could be extended and how it can be exploited to construct explicit lower bounds on distillation that take into account physical constraints.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d := \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmz \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmz} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend this to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmz} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmz &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n},
\end{align*}
to denote the phase space point for the composite system. 
For simplicity going forward, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\bmz \in \mathbb{Z}_d^{2n}$.

The displacement operators form the Heisenberg-Weyl group~\cite{Folland_1989, Bengtsson_2006} under matrix multiplication modulo phases,
\begin{equation}\label{eq:gp}
    {\rm{HW}}_d^n \coloneqq \{ \tau^k D_{\bmz}: k \in \mathbb{Z}_d, \bmz \in \cal{P}_d^n\}.
\end{equation}
The Clifford group $ \cal{C}_d^n $ is then defined as the set of unitaries that normalise the Heisenberg-Weyl group~\cite{Appleby_2005}. We may define the pure stabiliser states as those states obtained by acting on $|0\>$ with Clifford unitaries~\cite{cit:gross3} and, finally, the full set of stabilizer states as the convex hull of all pure stabilisers, namely all probabilistic mixtures of states of the form $U|0\>\<0|U^\dagger$ where $U$ is Clifford.

@@

In order to provide a complete decomposition of arbitrary quantum states and quantum operations we now define a basis of Hermitian observables that transforms covariantly under the action of the Clifford group. At every point $\bmz \in \P_d$ we define the phase-point operator
\begin{equation}\label{eq:ax}
	A_{\bmz} \coloneqq \frac{1}{d} \sum_{\bmy \in \cal{P}_d} \omega^{\eta(\bmz, \bmy)} D_{\bmy}, 
\end{equation}
where $\eta(\bmz, \bmz)$ is the symplectic inner product between any two points $\bmz,\z \in \P_d$, and is given explicitly by
\begin{equation}
	\eta(\bmz, \bmy) \coloneqq \bmy^T \begin{pmatrix}
		0  & \id \\ %\mathbbm{O}_n  & \id_n \\
		-\id & 0 \\ %-\id_n & \mathbbm{O}_n \\
	\end{pmatrix} \bmz,
\end{equation}
where $0, \id$ denote the $n\times n$ zero and identity matrices.

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\cal{H}_d)$ can be expressed as a linear combination of them, $\rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz}$, where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmz]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmz} \rho].
\end{equation}

@@

\subsection{Magic theories for quantum computation}
\label{sec:mono}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Probing magic protocols with reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the noise parameters $\epsilon, \epsilon'$, and free energies $F, F'$. In addition, a quantity $\phi$ (Equation~\ref{eq:phi}) relates the computational and energy bases, and corresponds to the degree to which the Wigner negativity of $\rho$ can be given a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}
A number of magic theories exist, where magic states are computational resource states with respect to a natural class of quantum operations that are considered free~\cite{Gour_2019}. One very natural class of free operations are those obtained from Clifford unitaries, Pauli measurements and the ability to discard quantum systems. However, there are also other candidates~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

@@

In this work we develop bounds that apply to any reasonable magic theory. The central idea is to apply majorization theory to the quasi-probability distributions that arise for magic states. We begin in the next section by explaining how majorization relates to the free operations in a theory of magic.

@@

Within the Wigner representation for odd prime dimension, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:veitch}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory are required to be a subset of
\begin{equation}
    \F \coloneqq \{ \rho: \W[\bmz]{\rho} \geq 0 \text{ for all } \bmz \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed quantum operations, i.e. the ones that are considered free within the theory. The most basic assumption we require on free operations is that they send any free state to another free state, meaning that a resource state cannot be created freely.

Any magic state protocol will correspond to a quantum channel $\E$, and so from Equation~(\ref{Wigner-Choi}) admits a Wigner representation $W_{\E}( \y | \bmz)$ that acts as a transition matrix mapping phase space points $\bmz \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\bmz \in \P_d} W_{\E}( \y | \bmz) W_\rho(\bmz),
\end{equation}
for any $\E$ and $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that if $\E$ is a free operation then the associated Wigner representation $W_{\E}(\y |\bmz)$ must form a stochastic matrix. In particular, all stabilizer operations correspond to stochastic matrices in the Wigner representation.

@@

Majorisation~\cite{cit:marshall, Blackwell_1953} is a collection of powerful tools that has found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations, and in its simplest form defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. A well-known result~\cite{cit:marshall} tells us that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be checked efficiently.

There is a natural generalisation, which is called $\bmd$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization~\cite{cit:horodecki2013}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as $\p \succ_{\r} \q$, if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, this majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from $\bmd$-majorization is also easily computed and again corresponds to checking a small number of inequalities.

@@

With this theory in place, we now turn to magic distillation protocols.

@@

We can now consider how physical constraints affect magic distillation rates starting in this section with an abstract setting based on the fixed-point structure of the physical operations accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure. This allows the use of $\mathbf{d}$--majorization, however in~\cref{sec:stab}, we drop the requirement of fixed-points and instead use the full relative majorization, before giving general entropic constraints in Section~\ref{section:entropies}.

 We begin by defining the following sub-theory of any magic theory $\R$.
\begin{definition}\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, we define the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those free operations that leave the distinguished state $\sigma \in \F$ invariant.
\end{definition}
This gives a simple way to break up any theory into smaller, more manageable parts, and where the union over all sub-theories returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho_1 \rightarrow \rho_2$ is possible in $\R$ if and only if the transformation $\rho_1 \rightarrow \rho_2$ is possible in at least one sub-theory $\R_\sigma$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
   Suppose the interconversion is possible in a $\R_\sigma$ via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer Fixed-Point theorem~\cite{cit:brouwer} this mapping must therefore have a fixed point $\sigma \in \F$. Therefore, $\E \in \O_\sigma$ and the interconversion is possible in $\R_\sigma$.
\end{proof}
We can compare with resource monotones, which are also called resource measures. A complete set of measures $\{\M_i\}_i$ is such that $\rho \rightarrow \sigma$ if and only if $\M_i(\rho) \ge \M_i(\sigma)$ for all $i$. The above set of sub-theories can therefore be viewed as a complete set of ``co-measures'' for the theory, where the parent resource theory's complex pre-order of states is mapped to a simpler pre-order for a particular $\R_\sigma$. In the next section we show that a given $\R_\sigma$ can be approximated by a majorization pre-order, and from this we can compute magic distillation bounds.

\subsection{Majorisation of Wigner distributions with fixed-point structure}
\label{sec:major_frag}
We now look at how fixed-points give constraints on magic protocols.
We consider some $\sigma \in \F$, and its corresponding $\R_\sigma$. We are interested in the ability to transform many copies of some noisy magic state $\rho$ towards a more pure form of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\bmz) < 0$ for some regions of $\bmz \in \P_d$. The state $\sigma$ is assumed to have a Wigner distribution with $W_\sigma(\bmz) > 0$ for all $\bmz$ in the phase space\footnote{A non-full rank state $\sigma$ can be handled as a limiting case in which we first add an infinitesimal fraction of depolarising noise $\epsilon (\id/d)$ and then take $\epsilon \rightarrow 0$ in the end.}.

@@

  Note, however, that this is not a single numerical constraint, but a family of constraints. For $n$ copies of a qudit system of dimension $d$ the number of terms in $W_{\rho}$ is $d^{2n}$, so imposing the Lorenz curve condition corresponds to \emph{exponentially} many constraints. However,  before computing explicit examples, in the next section we describe how these majorization constraints relate to other known monotones.

>>>>

\title{Constraints on magic state protocols from the statistical mechanics of \\Wigner negativity}

@@

To address this, concrete distillation protocols have been developed, such as the Bravyi-Haah qubit protocol that provides a quadratic reduction in noise per cycle~\cite{Bravyi2012}. Such distillation rates have been improved in more recent works~\cite{Jones_2013, haah2017magic, Hastings2018, Litinski_2019} and protocols for qudits of odd dimension have also been proposed~\cite{CampbellAnwar_2012, Anwar_2012, Dawkins_2015, Krishna2019, cit:prakash}. There is also analysis of magic protocols from the perspective of magic monotones, which provide upper bounds on distillation rates and find application in the analysis of gate synthesis~\cite{Campbell_2017, Howard_2017, Prakash_2018, Seddon_2021,leone2021renyi}. These frameworks for magic view magic states as resource states with respect to a natural class of quantum operations that are considered cheap, or free, such as stabilizer operations~\cite{Gour_2019, cit:ahmadi, cit:seddon, Wang_2019}. Beyond quantum computing, recent work has also looked at how magic can be used to analyse many-body physics~\cite{Sarkar_2020} and conformal field theories~\cite{White_2021} as well its connection to contextuality~\cite{Vega_2017, cit:howard2, Zurel_2020, Okay_2021}.

Work towards fault-tolerance is increasingly bridging the gap between abstract theory and experiment. Extensive work is being done on error mitigation~\cite{jones_2012, Li_2017, Temme_2017, Endo_2018, McClean_2017}, and the incorporation of hardware physics into the theoretical models~\cite{Kandala_2019, holmes_resource_2019, Colless_2018, song2018quantum, Bravyi_2021}. For example, the XZZX code~\cite{bonilla_ataides_xzzx_2021} is a variant of the surface code that incorporates noise bias explicitly and has been shown to attain the hashing bound of random codes~\cite{Bennett_1996}.

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated in the distillation bounds. This is achieved by considering how a given magic state protocol transforms a \emph{pair} of quantum states -- one being a noisy magic state, and the other a stabilizer state that is distinguished by the physics of the system (for example a state at a characteristic temperature, or the maximally mixed state) and acts as a reference state for the protocol. The magic distillation bounds can then be expressed in terms of the physics of the reference state (e.g.~free energy changes in the case of temperature). Such bounds are of potential interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols~\cite{Tuckett_2019, Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in key ways. We obtain distillation upper bounds without the use of monotones, but instead use tools from statistical mechanics and recent work in single-shot resource theories~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. Our analysis relies on the discrete Wigner representation of quantum systems, in which all states and operations can be described on a discrete phase space~\cite{Ferrie_2008, Okay_2021}. Crucially, we focus on magic states with negativity in their Wigner representation, which is known to be a necessary condition for universality in the state-injection model~\cite{cit:veitch, cit:mari, cit:gottesman, cit:knill, Campbell_2011}. However, taking a statistical mechanical perspective in this context raises a problem: the standard Boltzmann entropy is not defined for quasi-distributions. We circumvent this obstacle by making use of a more fundamental tool -- majorization theory~\cite{cit:marshall, Veinott_1971, Ruch_1976}. To our knowledge majorization of quasi-distributions has not been considered in quantum physics before, and therefore these methods could find application in other studies of non-classicality in quantum systems~\cite{Fine_1982, Allahverdyan_2018, arvidsson_2020, halpern_2018, Lostaglio_2018, Levy_2020}.

@@

We find that the analysis in the presence of negativity displays a range of features that do not appear in classical statistical mechanics, and leads to a picture of Wigner negativity in a quantum circuit being described as non-classical free energy that is processed under stochastic dynamics. This is demonstrated by non-monotonic Lorenz curves (Section~\ref{sec:lc}), and by single-shot entropies for general Wigner quasi-distributions that are well-defined, obey data-processing under stabilizer operations and can take on negative values (Section~\ref{section:entropies}).

@@

We show that the mana of a quantum state arises as a residue in the $\alpha$--parameter for R\'{e}nyi entropies, and gives the rate of divergence to $-\infty$ as we approach the Shannon entropy of a Wigner distribution (Equation (\ref{eq:mana-divergence})).
Finally, in Section~\ref{sec:lower_bounds} we discuss how our approach could be extended to construct explicit lower bounds on distillation that take into account physical constraints.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d := \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmz \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmz} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend this to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmz} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmz &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n},
\end{align*}
to denote the phase space point for the composite system. 
For simplicity, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\bmz \in \mathbb{Z}_d^{2n}$.

The displacement operators form the Heisenberg-Weyl group~\cite{Folland_1989, Bengtsson_2006} under matrix multiplication modulo phases,
\begin{equation}\label{eq:gp}
    {\rm{HW}}_d^n \coloneqq \{ \tau^k D_{\bmz}: k \in \mathbb{Z}_d, \bmz \in \cal{P}_d^n\}.
\end{equation}
The Clifford group $ \cal{C}_d^n $ is then defined as the set of unitaries that normalise the Heisenberg-Weyl group~\cite{Appleby_2005}. We may define the pure stabilizer states as those states obtained by acting on $|0\>$ with Clifford unitaries~\cite{cit:gross3} and, finally, the full set of stabilizer states as the convex hull of all pure stabilizers, namely all probabilistic mixtures of states of the form $U|0\>\<0|U^\dagger$ where $U$ is Clifford.

@@

In order to provide a complete decomposition of arbitrary quantum states and quantum operations we now define a basis of Hermitian observables that transforms covariantly under the action of the Clifford group. At every point $\bmz \in \P_d$ we define the phase-point operator
\begin{equation}\label{eq:ax}
	A_{\bmz} \coloneqq \frac{1}{d} \sum_{\bmy \in \cal{P}_d} \omega^{\eta(\bmz, \bmy)} D_{\bmy}, 
\end{equation}
where $\eta(\bmz, \y)$ is the symplectic inner product between any two points $\bmz,\y \in \P_d$, and is given explicitly by
\begin{equation}
	\eta(\bmz, \bmy) \coloneqq \bmy^T \begin{pmatrix}
		0  & \id \\ %\mathbbm{O}_n  & \id_n \\
		-\id & 0 \\ %-\id_n & \mathbbm{O}_n \\
	\end{pmatrix} \bmz,
\end{equation}
where $0, \id$ denote the $n\times n$ zero and identity matrices.

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho$ can be expressed as a linear combination of them, $\rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz}$, where the coefficients $\W{\rho}(\z)$ give the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmz]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmz} \rho].
\end{equation}

@@

\subsection{Magic theories for quantum computation}
\label{sec:mono}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Probing magic protocols with reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the noise parameters $\epsilon, \epsilon'$, and free energies $F, F'$. In addition, a quantity $\phi$ (Equation~\ref{eq:phi}) relates the computational and energy bases, and corresponds to the degree to which the Wigner negativity of $\rho$ can be given a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}
A number of magic theories exist, where magic states are computational resource states with respect to a class of quantum operations that are considered free~\cite{Gour_2019}. One natural class of free operations are those obtained from Clifford unitaries, Pauli measurements and the ability to discard quantum systems. However, there are also other candidates~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

@@

In this work we develop bounds that apply to any reasonable magic theory. The central idea is to apply majorization theory to the quasi-probability distributions that arise for magic states. In the next section we explain how majorization relates to the free operations in a theory of magic.

@@

Within the Wigner representation for odd prime dimension, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:veitch}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory are required to be a subset of
\begin{equation}
    \F \coloneqq \{ \rho: \W[\bmz]{\rho} \geq 0 \text{ for all } \bmz \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of free quantum operations. The most basic assumption we require on free operations is that they send any free state to another free state.

Any magic state protocol will correspond to a quantum channel $\E$, and so from Equation~(\ref{Wigner-Choi}) admits a Wigner representation $W_{\E}( \y | \bmz)$ that acts as a transition matrix mapping phase space points $\bmz \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\bmz \in \P_d} W_{\E}( \y | \bmz) W_\rho(\bmz),
\end{equation}
for any $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that if $\E$ is a free operation then the associated Wigner representation $W_{\E}(\y |\bmz)$ must form a stochastic matrix. In particular, all stabilizer operations correspond to stochastic matrices in the Wigner representation.

@@

Majorisation~\cite{cit:marshall, Blackwell_1953} is a collection of powerful tools that has found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations, and in its simplest form defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. It can be shown~\cite{cit:marshall} that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be checked efficiently.

There is a natural generalisation, which is called $\bmd$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization~\cite{cit:horodecki2013}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as $\p \succ_{\r} \q$, if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. The original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$.

@@

With this theory in place, we turn to magic distillation protocols.

@@

We can now consider how majorization constrains magic distillation rates starting in this section with an abstract setting based on the fixed-point structure of the physical operations accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure. This allows the use of $\mathbf{d}$--majorization, however in~\cref{sec:stab}, we drop the requirement of fixed-points and instead use the full relative majorization, before giving general entropic constraints in Section~\ref{section:entropies}.

 We begin by defining the following sub-theory of any magic theory $\R$.
\begin{definition}\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, we define the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those free operations that leave the distinguished state $\sigma \in \F$ invariant.
\end{definition}
This gives a simple way to break up any theory into smaller, more manageable parts, and where the union over all sub-theories returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho_1 \rightarrow \rho_2$ is possible in $\R$ if and only if the transformation $\rho_1 \rightarrow \rho_2$ is possible in at least one sub-theory $\R_\sigma$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
   Suppose the interconversion is possible in a $\R_\sigma$ via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer Fixed-Point theorem~\cite{cit:brouwer} this mapping must therefore have a fixed point $\sigma \in \F$. Therefore, $\E \in \O_\sigma$ and the interconversion is possible in $\R_\sigma$.
\end{proof}
We can compare with resource monotones, which are also called resource measures. A complete set of measures $\{\M_i\}_i$ is such that $\rho \rightarrow \sigma$ if and only if $\M_i(\rho) \ge \M_i(\sigma)$ for all $i$. The above set of sub-theories can therefore be viewed as a complete set of ``co-measures'' for the theory, where the parent resource theory's complex pre-order of states is mapped to a simpler pre-order for a particular $\R_\sigma$. In the next section we show that a given $\R_\sigma$ can be approximated by a majorization pre-order, and from this we compute magic distillation bounds.

\subsection{Majorisation of Wigner distributions with fixed-point structure}
\label{sec:major_frag}
We consider some $\sigma \in \F$, and its corresponding $\R_\sigma$. We are interested in the ability to transform many copies of some noisy magic state $\rho$ towards a more pure form of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\bmz) < 0$ for some regions of $\bmz \in \P_d$. The state $\sigma$ is assumed to have a Wigner distribution with $W_\sigma(\bmz) > 0$ for all $\bmz$ in the phase space\footnote{A non-full rank state $\sigma$ can be handled as a limiting case in which we first add an infinitesimal fraction of depolarising noise $\epsilon (\id/d)$ and then take $\epsilon \rightarrow 0$ in the end.}.

@@

  Note, however, that this is not a single numerical constraint, but a family of constraints. For $n$ copies of a qudit system of dimension $d$ the number of terms in $W_{\rho}$ is $d^{2n}$, so imposing the Lorenz curve condition corresponds to \emph{exponentially} many constraints.  Before computing explicit examples, in the next section we describe how these majorization constraints relate to other known monotones.
